The central objective of this proposal is to \textbf{facilitate the adoption of
machine learning for neuroscience experimental control}. Our approach builds on
Bonsai.ML, an extension of the widely used Bonsai platform, which
equips neuroscientists with accessible, high-performance tools for intelligent
and adaptive experimentation.

To achieve this aim, we will pursue three interconnected objectives. First, we
will promote \textbf{community uptake} by producing clear, comprehensive
documentation, delivering a Bonsai course, and disseminating Bonsai.ML through
a paper and a Bonsai developers conference. These activities will lower
barriers for non-specialist users, empower experimentalists to explore machine
learning in their own work, and foster a vibrant, self-sustaining community of
users and contributors.

Second, we will improve \textbf{software maintenance} by simplifying the
existing codebase and strengthening extensibility by integrating into Bonsai.ML
the C\# probabilistic programming language Infer.NET.  This will make Bonsai.ML
faster, more robust, easier to maintain, and more adaptable to evolving
research needs.  

Finally, we will ensure the \textbf{long-term sustainability and alignment} of
Bonsai.ML with the neuroscience community by convening an internationally
recognised steering committee. This group will provide guidance on priorities,
governance, and strategic planning, ensuring that Bonsai.ML remains a reliable,
community-driven platform for cutting-edge experimental neuroscience.  

