\paragraph{Staff costs:} cover the salary of two research software engineers working on the project.

\paragraph{Community engagement:} The directly incurred \texttt{Other costs} cover the costs of one training curse and one developers conference. These cost only cover food expenses, as the venue is provided free of charge by the Sainsbury Wellcome Centre.

\paragraph{Cash contributions:}
NeuroGEARS will contribute £5,733.60 in cash towards salary support for one RSE.

\paragraph{In-kind contributions:}
NeuroGEARS will provide in-kind contributions valued at £100,000, including:

\begin{enumerate}
    \item Supporting the organisation of a Bonsai course, including a Bonsai.ML module, and providing instructors for its delivery.

    \item Supporting the organisation of the 2026 Bonsai Development Conference, including a Bonsai.ML session, and providing instructors for its delivery.

    \item Organising and deploying a Bonsai booth, including a Bonsai.ML exhibit, at the 2025 Society for Neuroscience Annual Meeting, similar to that deployed in 2024.

    \item Delivering Bonsai and Bonsai.ML lectures at the annual
    \href{https://github.com/joacorapela/statNeuro2025}{Sainsbury Wellcome Centre Statistical Neuroscience course}.

    \item Contributing director time to the Bonsai.ML weekly meetings.

    \item Providing Bonsai technical support to members of the Bonsai.ML team.
\end{enumerate}

\subsection{Total funding requested:} £283,384.02

\subsection{Directly incurred - Staff:} £130,539.22

\subsection{Directly incurred - Travel and Subsistence:} £0.0

\subsection{Directly incurred - Other:} £13,600

\subsection{Directly allocated - Staff:} £0.0

\subsection{Directly allocated - Estates:} £26,100.80

\subsection{Directly allocated - Other:} £0.0

\subsection{Indirects:} £113.144

\subsection{Exceptions:} £0.0

