\documentclass[12pt]{article}

\usepackage[colorlinks=true]{hyperref}
\usepackage{xcolor}

\usepackage{tikz}

\makeatletter
\newcommand*{\radiobutton}{%
  \@ifstar{\@radiobutton0}{\@radiobutton1}%
}
\newcommand*{\@radiobutton}[1]{%
  \begin{tikzpicture}
    \pgfmathsetlengthmacro\radius{height("X")/2}
    \draw[radius=\radius] circle;
    \ifcase#1 \fill[radius=.6*\radius] circle;\fi
  \end{tikzpicture}%
}
\makeatother

\usepackage{enumitem,amssymb}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

\title{Letter of Intent for \href{https://www.software.ac.uk/research-software-maintenance-fund/round-1}{Research Software Maintenance Fund -- Round 1}}
\author{}

\newenvironment{instruction}{\par\color{red}}{\par}

\begin{document}
\maketitle

\section{Application name}

Consolidating Bonsai as the Standard for Neuroscience Intelligent Experimental Control

\section{Size of grant}

\begin{itemize}
\item[\radiobutton] Small award (up to 1 year / £150k)
\item[\radiobutton*] Large Award (up to 2 years / £500k)
\end{itemize}

\section{Proposal summary / scope of work (500 words)}

\begin{instruction}

Please provide a short summary of the work being proposed.

\end{instruction}

\input{summary}

\section{Categories of work}

\begin{todolist}
	\item Technical
	\item[\done] Documentation
	\item[\done] Training
	\item[\done] Community
    \item[\done] Governance
\end{todolist}

\section{Project team (250 words)}

\begin{instruction}

Briefly describe why your team is the appropriate one to carry out these
    activities. You will be asked to provide details of the Project Lead and
    Co-Leads on the Team Details page, along with their connection to the
    software.

\end{instruction}

\input{projectTeam}

\section{Benefit to UK research (250 words)}

\begin{instruction}

    Briefly describe the expected value of the proposed work to UK research.
    You may wish to describe the fields of research or types of research method
    the software is used in, and describe the benefits to a particular
    community. You may also want to give examples of what research is enabled
    by the software.


\end{instruction}

\section{Landscape analysis (250 words)}

\begin{instruction}

    Briefly describe the other software (either proprietary or open source)
    that is primarily used by the research community addressed by the work in
    this proposal. Summarise, to the best of your knowledge, how the software
    in this proposal compares to these other software in terms of user base
    size, usage, and maturity. Describe, if appropriate, how the other software
    interacts with the software in this proposal.

\end{instruction}

\section{Measure of impact (optional) (250 words)}

\begin{instruction}

    Briefly describe how you would measure the impact of the proposed work. You
    may have existing measures that you use, or you may propose new measures.
    For the Expression of Interest we are looking to understand how you would
    approach this, and what you think is important to measure. In the full
    application, you will be asked to define specific measures.

\end{instruction}

\end{document}
