
The large field of technologies serving experimental control and behaviour
monitoring is traditionally occupied either by domain-specific graphical user
interfaces for control and recording of specific devices and experiment types
(e.g.\ Open Ephys GUI\footnote[16]{\url{open-ephys.org/gui/}}, Miniscope DAQ
Software\footnote[17]{\url{github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}}) or by
real-time control frameworks for specifying task logic using state machine or
similar formalisms (e.g. NIMH ML\footnote[18]{\url{monkeylogic.nimh.nih.gov/}},
pyControl\footnote[19]{\url{pycontrol.readthedocs.io/en/latest/}},
Autopilot\footnote[20]{\url{docs.auto-pi-lot.com/en/latest/}},
Sanworks\footnote[21]{\url{sanworks.io/index.php}}).

These dedicated interfaces are typically very comfortable for experimenters in
the specific domain for which the tool is designed, but can become unwieldy
with the introduction of a new device or task variation from outside their
usual scope.
%
Alternatively, one can use a more general programming language such as Python
or MATLAB, with the disadvantage of the code being harder to understand,
maintain, and change.

Programming languages like LabVIEW straddle the middle ground and provide a
high-level, flexible visual interface for composing data acquisition and
control systems. Unlike Bonsai, however, the graphical elements of LabVIEW are
heterogeneous and very fine grained, thus requiring long and complex logical
structures to implement even a simple experimental control system.
%
A severe disadvantage of LabVIEW for the research community is that it is not
free nor open source.

By providing an extremely simple, yet flexible visual syntax, Bonsai provides
the opportunity for even complete non-programmers to design and successfully
customise relatively complex experiments from the ground up. It is this
capability in particular which has made Bonsai such an attractive standard tool
in experimental neuroscience.
%
In addition, Bonsai is free and open source.

