The consolidation of Bonsai as a standard for intelligent experimental control
requires expertise in software engineering, machine learning and experimental
neuroscience.
%
In addition, Bonsai is a very unique programming language. It is a visual
programming language, using the reactive programming framework and implemented
in C\#.
%
Therefore, a very unique set of software development skills are required to
integrate ML functionality in Bonsai.

We are a very cohesive team, with different levels of seniority, and with
complementary expertise to successfully deliver the proposed software
maintenance tasks.

\begin{description}

    \item[Prof.~Maneesh Sahani and Prof.~Thomas Mrsic-Flogel] lead the Gatsby
        Computational Neuroscience Unit and the Sainsbury Wellcome Centre,
        first-rate research centers on computational and experimental
        neuroscience, respectively. As projects lead and co-lead in the BBSRC
        grant that funded the creation of Bonsai.ML, they are deeply involved
        on its success.

    \item[Dr.~Gonçalo Lopes] is the creator and main developer of Bonsai, he
        knows every detail of it, and he is heavily involved in the Bonsai.ML
        package.
        %
        NeuroGEARS, the non-profit company directed by Dr.~Lopes that is the
        main contributor to the development of Bonsai, is a business partner in
        the grant that funded the creation of Bonsai.ML.
        %
        Dr.~Lopes meets weekly with Dr.~Rapela and Dr.~Guilbeault, developers
        of Bonsai.ML, to brainstorm about the progress of the project.

    \item[Dr.~Joaquín Rapela] has ample experience in signal processing,
        machine learning, neuroscience, neural data analysis and software
        development. He leads the implementation of the Bonsai.ML package.

    \item[Dr.~Nicholas Guilbeault] combines unique training in biology and
        neuroscience with large Bonsai software development. He created the
        BonZeb Bonsai
        package\footnote{\url{https://github.com/ncguilbeault/BonZeb}} and he
        is the core developer of the Bonsai.ML package, including its
        documentation.

\end{description}

